---
title: Visualizing Data from The nycflights13 Package
author: "Mariia Karnaukh"
date: "2024-01-31"
output: html_document
---

# Visualizing Flights of the New York Airport in 2013.

The following analysis makes use of the `nycflights13` package that contains several data-sets on airlines, airports, flights, planes and weather for the New York airport in the year 2013. I prepare below some visualization to better understand the trends presented in these data-sets.

## Step 0:

Loading the nycflights13 package and check what data-sets are bundled.

```{r}
library(nycflights13)
data(package = 'nycflights13')
```

## Step 1:

Visualizing the distribution of arrival delays per origin airport.

```{r}
library(ggplot2)
```

```{r, echo=FALSE}
ggplot(flights, aes(origin,arr_delay)) + geom_boxplot() +
  theme_grey() +
  labs(title = "Distribution of Arrival Delays per Origin Airport",
                                                           x = "Origin Airport",
                                                           y = "Arrival Delay (minutes)")
```

## Step 2:

Visualizing the distribution of arrival delays per destination airport.

```{r out.width = '100%'}
ggplot(flights, aes(dest,arr_delay, label = dest)) + 
  geom_boxplot(outlier.size = 0.5, lwd = 0.5) + 
  theme_grey() + 
  labs(title = 'Flights Departing from NY in 2013', 
       x = 'Destination Airport', 
       y = 'Arrival Delays (Minutes)') + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 4), 
        axis.title.x = element_text(size = 9), 
        axis.title.y = element_text(size = 9))
```

## Step 3:

Computing and visualizing the average arrival delay per destination.

```{r out.width = '100%'}
library(dplyr)
num_NA <- summary(is.na(flights))
num_NA ##arr_delay :9430 NA, about 3% of the data, drop NA

flights_c <- flights %>%
  filter(!(is.na(arr_delay)))

num_NA2 <- summary(is.na(flights_c))

avg_delay_table <- flights_c %>%
  group_by(dest) %>%
  summarise(avg_delay_dest = mean(arr_delay))
print(avg_delay_table)

num_NA3 <- summary(is.na(avg_delay_table))

ggplot(avg_delay_table, aes(reorder(dest, avg_delay_dest),avg_delay_dest, label = dest)) +
  geom_col() + labs(title = 'Flights Departing from NY in 2013',
                    x = 'Destination Airport',
                    y = 'Average Delay (Minutes)') +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 4),
        axis.title.x = element_text(size = 9),
        axis.title.y = element_text(size = 9))
```

## Step 4:

Redoing the above plot by showing the actual name of the destination airport instead of it's FAA id.

```{r out.width = '100%'}
data(airports)

avg_delay_table2 <- avg_delay_table %>%
  rename(faa = dest)

table2 <- avg_delay_table2 %>%
  left_join(airports, by = 'faa')

#Check for NA
num_NA4 <- summary(is.na(table2))

# Drop NA :100 unique values
table2 <- table2 %>%
  filter(!(is.na(lat)))

#Graph
ggplot(table2,
       aes(reorder(name, avg_delay_dest),avg_delay_dest, label = name)) +
  geom_col() + labs(title = 'Flights Departing from NY in 2013',
                    x = 'Destination Airport',
                    y = 'Average Delay (Minutes)') +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 4),
        axis.title.x = element_text(size = 9),
        axis.title.y = element_text(size = 9))
```

## Step 5:

Coloring the bars by the timezone of the airport.

```{r out.width = '100%'}
ggplot(table2,
       aes(reorder(name, avg_delay_dest),avg_delay_dest,
           label = name, fill = tzone)) +
  geom_col() +
  labs(title = 'Flights Departing from NY in 2013',
                    x = 'Destination Airport',
                    y = 'Average Delay (Minutes)',
                    fill = 'Timezone') +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, size = 4),
        axis.title.x = element_text(size = 9),
        axis.title.y = element_text(size = 9),
        legend.position = 'top',
        legend.direction = 'horizontal') +
  guides(fill = guide_legend(nrow = 1))
```

## Step 6:

Visualizing destination airports on a world-map with the point sizes relative to the number of flights to there from NY.

```{r out.width = '100%'}
# Add a new column with the number of time the airport was flown to
count_by_dest <- flights %>%
  count(dest, name = 'count_by_dest')
count_by_dest
airports_with_faa_count <- left_join(airports, count_by_dest, by = c('faa' = 'dest'))
airports_with_faa_count <- airports_with_faa_count %>%
  filter(!is.na(count_by_dest))
airports_with_faa_count

# Create the map using ggplot2
world <- map_data('world')

ggplot() +
  geom_polygon(data = world, aes(x = long, y = lat, group = group),
               fill = 'lightblue') +
  geom_point(data = airports_with_faa_count,
             aes(x = lon, y = lat, size = count_by_dest), color = 'red', alpha = 0.7) +
  scale_size_continuous(range = c(1, 5)) +
  labs(title = 'Flights Departing to NY in 2013', x = 'Longitude',
       y = 'Latitude', size = 'N') +
  theme_minimal()
```

## Step 7:

Computing the average departure and arrival delays, also the average air time and distance per destination, then passing the scaled dataframe through MDS to visualize the similarities/dissimilarities of the destination airports.

```{r out.width = '100%'}
avg_vars <- flights %>%
  group_by(dest) %>%
  summarize(avg_dep_delay = mean(dep_delay, na.rm = TRUE),
            avg_arr_delay = mean(arr_delay, na.rm = TRUE),
            avg_air_time = mean(air_time, na.rm = TRUE),
            avg_distance = mean(distance, na.rm = TRUE))
head(avg_vars)

dest_codes <- avg_vars$dest

avg_vars2 <- scale(avg_vars[, -1])
head(avg_vars2)

mds <- cmdscale(dist(avg_vars2))
head(mds)

#Plot using ggplot, convert to a data frame
mds <- as.data.frame(cmdscale(dist(scale(avg_vars2))))

#Flip horizontally
mds$V1 <- mds$V1 * -1
mds$dest <- dest_codes

# Plot MDS results using ggplot2
library(ggrepel)
ggplot(mds, aes(x = V1, y = V2, label = dest)) +
  geom_point() +
  geom_text_repel(aes(label = dest), size = 3, max.overlaps = 50) +
  labs(title = "MDS Visualization of Destination Airports",
       x = NULL,
       y = NULL) +
  theme_minimal()
```
